\LARGE \bf Exploiting Non-Slip Wall Contacts \\ to Position Two Particles Using The Same Control Input

### 2018-04-17 IROS17/theory2robAlgs.tex
Edits to the paper and pictures.

\begin{figure}
\centering
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
\caption{\label{fig:TwoRegions}
Boundary interaction is used to change the relative positions of the robots. Each robot gets the same control input. 
(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
}
\end{figure}
If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move. This may cause the algorithm not to return the shortest path, but the difference is not significant.

>>>>

\begin{figure}
\centering
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
\caption{\label{fig:TwoRegions}
Boundary interaction is used to change the relative positions of the particles. Each particle gets the same control input. 
(left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
}
\end{figure}
If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move. This may cause the algorithm not to return the shortest path, but the difference is not significant.

### 2018-04-17 journalWallFriction/SwarmShapeControl.tex
One algorithm added
A lot of new images
Edits to the paper and pictures.

% paper title
\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using a Shared Input}

% You will get a Paper-ID when submitting a pdf file to the conference system
\author{Shiva Shahrokhi, Jingang Shi, and Aaron T. Becker% <-this % stops a space
\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
}
%\affil{Department of Electrical and Computer Engineering, \\
% University of Houston, Houston, TX 77204-4005 USA\\
% {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}}
%\thanks{S. Shahrokhi, A. Mahadev and  A. Becker are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204-4005 USA {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}
%}
%} %\end thanks

@@

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
The walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
However, in vivo environments are usually not square. 
This work extends the previous work to convex workspaces and 3D positioning. 
This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.

>>>>

% paper title
\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using The Same Control Input}

% You will get a Paper-ID when submitting a pdf file to the conference system
\author{Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker% <-this % stops a space
\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
}
%\affil{Department of Electrical and Computer Engineering, \\
% University of Houston, Houston, TX 77204-4005 USA\\
% {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}}
%\thanks{S. Shahrokhi, A. Mahadev and  A. Becker are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204-4005 USA {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}
%}
%} %\end thanks

@@

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 This paper investigates control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
Because in vivo environments are usually not square, this work extends the previous work to convex workspaces including circles and 3D positioning. 
This paper also implements the algorithms using a hardware setup inspired by gastrointestinal tract.

### 2018-04-17 journalWallFriction/intro.tex
Edits to the paper and pictures.

\begin{figure}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\end{overpic}
\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\end{overpic}
\caption{\label{fig:IntroPic}
Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
} \vspace{-1em}
\end{figure}
%\todo{add the picture of magnetic setup}

@@

This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which consider only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.

>>>>

\begin{figure}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\put(28,-8){magnetic setup}\end{overpic}
\vspace{1em}
\caption{\label{fig:IntroPic}
Workspace and magnetic setup for an experiment of positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
} \vspace{-1em}
\end{figure}
%\todo{add the picture of magnetic setup}

@@

This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-04-17 journalWallFriction/theory.tex
Edits to the paper and pictures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Boundary Interaction Model}\label{subsec:WallFriction}

@@

 If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
 \begin{align}\label{eq:swarmDynamicsAndFric} 
\dot{\mathbf{x}}_i(t)
 &=
 \mathbf{u}(t)
 +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), ~i \in [1,n].\\
 F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
  - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
 \\
 0 & \textrm{else} 
 \end{cases}\nonumber
 \end{align}
 Here $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$, and
 $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary.

@@

 \subsection{Shortest Path}
 The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments shown in Fig.~\ref{fig:shortestpath}. The problem can be simplified by choosing the coordinate system carefully. We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point by a radius e and angle $\beta$, $E=e(\cos \beta,\sin \beta)$. Then define a symmetry point about S of line OP named T. Then the length of the two line segments is 
 \begin{equation}
\scriptsize{ \sqrt{(s-\cos \theta)^2+(-\sin \theta)^2}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}}
 \end{equation}
 which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to

@@

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.}
%\vspace{-1em}
}
\end{figure}

 The length of the two line segments as a function of $\theta$ is drawn in the right plot. There are several simple solutions. If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0. If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$. Label the origin $O$. The optimal solution shows that the angle $\angle{OPS}$ (from the origin to P to S) is the same as the angle $\angle{OPE}$ (from the origin to P to E). We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are S and E. When the ellipse is tangent to the circle, the point of tangency is exactly P. 
  Since the distance from the origin to P is always 1, we can set up three equalities using the law of sines:
 From triangle OSP: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{SP}$, and from triangle OEP: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{EP}$. If we mirror the point S about the $\theta$ axis and label this point C, from triangle CEO: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{CE}$.

 Simplifying this system of equations results in: $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi<=\beta<=\pi$. We evaluate each and select the solution that results in the shortest length path. Note that the optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM and RRT* are not suited \cite{lavalle2006planning,kavraki1996probabilistic}. Instead, this paper employs a graph search. This section starts with a boundary interaction model in subsection \ref{subsec:WallFriction}.

Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle. In subsections \ref{subsec:square} and \ref{subsec:circular} we provide shortest-path results for two representative workspaces, squares and circles.
\subsection{Boundary Interaction Model}\label{subsec:WallFriction}

@@

 If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
 \begin{align}\label{eq:swarmDynamicsAndFric} 
\dot{\mathbf{x}}_i(t)
 &=
 \mathbf{u}(t)
 +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), ~i \in [1,n].\\
 F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
  - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
 \\
 0 & \textrm{else}.
 \end{cases}\nonumber
 \end{align}
 Here  $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary, and
 $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$.

@@

 \subsection{Shortest Path in Square}\label{subsec:square}

 If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error  to zero, and  the third move takes the particles to their final positions. 
The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
 Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig. \ref{fig:reflection}.

\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
\caption{\label{fig:reflection}
The shortest path between two points in a square. Reflecting the goal position across the boundary wall and plotting a straight line gives the shortest path.
} \vspace{-1em}
\end{figure}

 \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
 The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments shown in Fig.~\ref{fig:shortestpath}. The problem can be simplified by choosing the coordinate system carefully. We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point $E$ by a radius $e$ and angle $\beta$, $E=e(\cos \beta,\sin \beta)$. Then the length of the two line segments is 
 \begin{equation}
\scriptsize{ \sqrt{(s-\cos \theta)^2+\sin^2 \theta}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}},
 \end{equation}
 which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to

@@

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right. See \cite{shortestPathMathematica}.}
%\vspace{-1em}
}
\end{figure}

 The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}. There are several simple solutions. If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0. If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$. Label the origin $O$. The optimal solution shows that the angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is exactly $P$. 
  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
 From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$. If we mirror the point $S$ about the $\theta$ axis and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.

 Simplifying this system of equations results in: $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$. We evaluate each and select the solution that results in the shortest length path. This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection. For an interactive mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.
